---
title: "ROP frequentist meta-analysis - Gestational age and birth weight as potential moderators"
author: "Aman Josan"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document: 
    self_contained: yes
    
    

editor_options: 
  chunk_output_type: inline
---




<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-right{
  float: right;
  width: 50%;
  text-align: right;
}
</style>





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, comment=NA, cache.rebuild = TRUE)
library(here)
library(dplyr)
library(readxl)
library(robumeta)
library(gemtc)
library(rjags)
library(reshape2)
library(kableExtra)
library(metafor)
library(dmetar)
library(tidyverse)
```


<br><br>


### Type 1 ROP
<br>

```{r, message=FALSE, warning=FALSE}
data <- suppressMessages(read_excel(("C:\\Users\\ajosan\\Desktop\\R_scripts\\ROP_meta_analysis\\data\\ROP data 01092021-modified.xlsx"),sheet = 2, range = "b3:ba41", col_names = T))

df <- data %>% select("study",
                      "treatment-Avastin",
                      "treatment-Lucentis",
                      "treatment-Eylea",
                      "treatment-laser",
                      "study design", 
                      "no. Eyes-Avastin",
                      "no. Eyes-Lucentis",
                      "no. Eyes-Eylea",
                      "no. Eyes-laser",
                      "no. retreated-Avastin",
                      "no. retreated-Lucentis",
                      "no. retreated-Eylea",
                      "no. retreated-laser",
                      "mean gestational age - Avastin (wks)",
                      "mean gestational age - Lucentis (wks)",
                      "mean gestational age - Eylea (wks)",
                      "mean gestational age - Laser (wks)",
                      "mean birth weight - Avastin (wks)",
                      "mean birth weight - Lucentis (wks)",
                      "mean birth weight - Eylea (wks)",
                      "mean birth weight - Laser (wks)"
                      ) 

df[df == "n/a" ] <- NA

df[,c(7:22)] <- suppressWarnings(lapply(df[,c(7:22)], as.numeric))



df$`n_responders-Avastin` <- 
  df$`no. Eyes-Avastin` - df$`no. retreated-Avastin`
df$`n_responders-Lucentis` <- 
  df$`no. Eyes-Lucentis` - df$`no. retreated-Lucentis`
df$`n_responders-Eylea` <- 
  df$`no. Eyes-Eylea` - df$`no. retreated-Eylea`
df$`n_responders-laser` <- 
  df$`no. Eyes-laser` - df$`no. retreated-laser`

df <- df[,-c(11:14)]

df <- mutate(df, study_id = 1:nrow(df)) # This adds a study id
df <- df[c(23,1:22)] # rearranges columns

df_table <- df[,-c(1)]
### print table of studies
df_table %>%
  kbl(caption = "All studies which considered type 1 ROP") %>%
  kable_classic(full_width = T, html_font = "Cambria", font_size=9) %>%
  kable_styling("striped") 

no_studies <- nrow(df_table)
no_patients <- sum(df[,c(8:11)], na.rm = T)
```

<br><br>

number of studies=`r no_studies`


Number of patients=`r no_patients`

<br><br>


## Event count data and conventional meta-analysis of proportions


<br>


### Transform data (Dichotomous data)

<br>

Here we define the number of responders as the number of patients treated minus the number of patients requiring re-treatment. Hence the measure of treatment effectiveness we are applying is the number of patients requiring no re-treatment.


Since we are dealing with raw prevalence (or event rate) data (i.e. sample size and number of treatment responders) rather than correlations (i.e. Pearsons r for treatment correlations) we need to perform a meta-analysis of proportions.  Most meta-analysis in the literature deal with correlations and p-values.



The proportion of treatment responders ($p$) is given by

$$\begin{align}
p = \frac{x}{n}
\end{align}$$

<br>

where $x$ is the number of treatment responders in a study and $n$ is the total number treated with that treatment in the study.



Again data is quite skewed (as is normally the case in number of event data), hence use double arcsine transformation for the conventional meta-analysis again.




```{r, results="hide"}
### prep dataframe for analysis by having t,n,r columns
df2 <- df

df2$t.1 <- 1 #"Avastin" 
df2$t.2 <- 2 #"Lucentis" 
df2$t.3 <- 3 #"Eylea"
df2$t.4 <- 4 #"Laser"

df2 <- df2[c(1,2,24:27,8:11,20:23,12:15,16:19)]

colnames(df2)[7] <- "n.1"
colnames(df2)[8] <- "n.2"
colnames(df2)[9] <- "n.3"
colnames(df2)[10] <- "n.4"
colnames(df2)[11] <- "r.1"
colnames(df2)[12] <- "r.2"
colnames(df2)[13] <- "r.3"
colnames(df2)[14] <- "r.4"
colnames(df2)[15] <- "ga.1"
colnames(df2)[16] <- "ga.2"
colnames(df2)[17] <- "ga.3"
colnames(df2)[18] <- "ga.4"
colnames(df2)[19] <- "bw.1"
colnames(df2)[20] <- "bw.2"
colnames(df2)[21] <- "bw.3"
colnames(df2)[22] <- "bw.4"


studyID <- df2[,c(1,2)]
df2 <- suppressWarnings(data.frame(lapply(df2,as.numeric)))
df2 <- df2[,-c(2)]
```

```{r, include=FALSE}
### reshape to group by treatment type
df_long <- reshape(df2, 
  varying = c("t.1","t.2","t.3","t.4",
              "n.1","n.2","n.3","n.4",
              "r.1","r.2","r.3","r.4",
              "ga.1","ga.2","ga.3","ga.4",
              "bw.1","bw.2","bw.3","bw.4"), 
  timevar = "treatment",
  direction = "long")

df_long <- df_long[,c(1,3,4,5,6,7)] %>% arrange(study_id)

df_long <- df_long[!is.na(df_long$r), ]
df_long <- df_long[!is.na(df_long$n), ]

colnames(df_long) <- c("study_id", "treatment", "sampleSize", "responders", "gest.age", "b.weight") 

df_long <- merge(x = df_long, y = studyID, by = "study_id", all.x = TRUE)
df_long <- df_long[c(1,7,2,3,4,5,6)]
df_long$treatment <- as.character(df_long$treatment)
df_long$trt <- ""
df_long$trt[df_long$treatment=="1"] <- "Avastin"
df_long$trt[df_long$treatment=="2"] <- "Lucentis"
df_long$trt[df_long$treatment=="3"] <- "Eylea"
df_long$trt[df_long$treatment=="4"] <- "Laser"
df_long <- df_long[c(1,2,3,8,4,5,6,7)]
```

```{r, include=FALSE, warning=FALSE}
### use metafor package to plot forest, funnel plots and - not available in gemtc
metafor_data <- df_long
colnames(metafor_data) <- c("study", "authors","treatment","trt","ni","xi","g.age","b.weight")
metafor_data <- mutate(metafor_data, id = 1:nrow(metafor_data))
metafor_all_data <- metafor_data[c(9,1,2,4,5,6,7,8)]



### use escalc to find effect size in terms of freeman & Tukey proportions
metafor_all_df <- metafor::escalc(data=metafor_all_data, measure="PFT", xi=xi, ni=ni, slab=paste(authors, trt), add = 0)


### perform mixed effects model on each group
metafor_all_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight,data=metafor_all_df, method = "REML")

metafor_Avastin_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight,data=metafor_all_df, method = "REML", subset=(trt=="Avastin"))

metafor_Lucentis_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight,data=metafor_all_df, method = "REML", subset=(trt=="Lucentis"))

#metafor_Eylea_res_mod <- 
#  metafor::rma(yi,vi,mods = ~ g.age+b.weight,data=metafor_all_df, method = "REML", subset=(trt=="Eylea"))

metafor_Laser_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight,data=metafor_all_df, method = "REML", subset=(trt=="Laser"))
```



```{r}
Avastin_stats <- 
  data.frame(capture.output(metafor_Avastin_res_mod)) %>%
  kbl(caption = "Avastin") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Avastin_stats


```


```{r}
Lucentis_stats <- data.frame(capture.output(metafor_Lucentis_res_mod)) %>%
  kbl(caption = "Lucentis") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Lucentis_stats

```


```{r eval=FALSE, include=FALSE}
Eylea_stats <- data.frame(capture.output(metafor_Eylea_res_mod)) %>%
  kbl(caption = "Eylea") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Eylea_stats
```


```{r}
Laser_stats <- 
  data.frame(capture.output(metafor_Laser_res_mod)) %>%
  kbl(caption = "Laser") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Laser_stats
```


<br><br>

Looking at the lines titled "test of moderators" and "g.age" and "b.weight" under "model results" we see that the model predictions find neither gestational age nor birth weight have a significant moderating factor. 


For example for those treated with laser, the p-value for the combined moderating effect of GA and BW is 0.9457 and with individual moderating effect p-values for GA=0.999 and BW=0.9161.



<br>

Hence all treatments now have non-significant correlations for gestational age and birth weight as moderators for effect size. Eylea moderator analysis not possible as too few samples.

<br>


Performing a frequentist meta-regression and generating a scatter plot with regression line and confidence intervals to see the relationship although now there is no significant correlation so these plots can be mostly ignored. Here I've only shown plots for gestational age as birth weight plots look pretty similar. I can plot them if needed but as they are non-significant I haven't bothered here.


<br><br>

Hence, for Lucentis (flat-ish line shows there is no real correlation to GA):

<div class = "row">
<div class="column-left">
```{r, warning=FALSE}
Lucentis_data <- metafor_all_data %>%
filter(grepl("Lucentis", `trt`)) 

Lucentis_data_mod <- 
            metafor::escalc(data = Lucentis_data,  
                            measure = "PFT", xi=xi, ni=ni, 
                            slab = paste(authors, trt), add = 0)

Lucentis_modplot <- metafor::rma(yi, vi, 
                            mods = ~ g.age,
                            data = Lucentis_data_mod,                                         
                            method = "REML")

Lucentis_predict <- 
          predict(Lucentis_modplot, newmods=c(20:35),     
          transf=transf.ipft.hm, 
          targ=list(ni=Lucentis_data_mod$ni))


wi=1/sqrt(Lucentis_data_mod$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(Lucentis_data_mod$g.age,transf.ipft.hm(Lucentis_data_mod$yi,targ=list(ni=Lucentis_data_mod$ni)),cex=size,pch=1,las=1,xlab="Gestational age (wks)", ylab="Proportional effect size")
lines(20:35,Lucentis_predict$pred,col="navy")
lines(20:35,Lucentis_predict$ci.lb,lty="dashed",col="maroon")
lines(20:35,Lucentis_predict$ci.ub,lty="dashed",col="maroon")
ids=c(1:18)
text(Lucentis_data_mod$g.age[ids],transf.ipft.hm(Lucentis_data_mod$yi,targ=list(ni=Lucentis_data_mod$ni))[ids],ids,cex=0.8,pos=1)
```

</div>
<div class="column-right">

```{r}
Lucentis_data_table <- Lucentis_data_mod[c(3,6,7,8)]
Lucentis_data_table <- mutate(Lucentis_data_table, 
                        study = 1:nrow(Lucentis_data_table))
Lucentis_data_table <- Lucentis_data_table[,c(5,1,2,3,4)]
colnames(Lucentis_data_table) <- c("Study", "Authors", "Sample size", "Gestational age","Birth weight")

Luc_study_table <- 
  Lucentis_data_table %>%
  kbl(caption = "Lucentis studies and assigned ID's") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=9)

Luc_study_table

no_Luc_studies <- nrow(Lucentis_data_table)
no_Luc_patients <- sum(Lucentis_data_table[,c(3)], na.rm = T)
```
</div>
</div>


<br>

N.B. Larger circles represent larger studies. 


<br><br>


Scatterplot for Avastin again showing a flat-ish line representing no real correlation


<div class = "row">
<div class="column-left">
```{r, warning=FALSE}
Avastin_data <- metafor_all_data %>%
filter(grepl("Avastin", `trt`)) 

Avastin_data_mod <- 
            metafor::escalc(data = Avastin_data,  
                            measure = "PFT", xi=xi, ni=ni, 
                            slab = paste(authors, trt), add = 0)

Avastin_modplot <- metafor::rma(yi, vi, 
                            mods = ~ g.age,
                            data = Avastin_data_mod,                                          
                            method = "REML")

Avastin_predict <- 
          predict(Avastin_modplot, newmods=c(20:35),     
          transf=transf.ipft.hm, 
          targ=list(ni=Avastin_data_mod$ni))


wi=1/sqrt(Avastin_data_mod$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(Avastin_data_mod$g.age,transf.ipft.hm(Avastin_data_mod$yi,targ=list(ni=Avastin_data_mod$ni)),cex=size,pch=1,las=1,xlab="Gestational age (wks)", ylab="Proportional effect size")
lines(20:35,Avastin_predict$pred,col="navy")
lines(20:35,Avastin_predict$ci.lb,lty="dashed",col="maroon")
lines(20:35,Avastin_predict$ci.ub,lty="dashed",col="maroon")
ids=c(1:18)
text(Avastin_data_mod$g.age[ids],transf.ipft.hm(Avastin_data_mod$yi,targ=list(ni=Avastin_data_mod$ni))[ids],ids,cex=0.8,pos=1)
```

</div>
<div class="column-right">

```{r}
Avastin_data_table <- Avastin_data_mod[c(3,6,7,8)]
Avastin_data_table <- mutate(Avastin_data_table, 
                        study = 1:nrow(Avastin_data_table))
Avastin_data_table <- Avastin_data_table[,c(5,1,2,3,4)]
colnames(Avastin_data_table) <- c("Study", "Authors", "Sample size", "Gestational age", "Birth weight")

Avastin_study_table <- 
  Avastin_data_table %>%
  kbl(caption = "Avastin") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=9)

Avastin_study_table

no_Avastin_studies <- nrow(Avastin_data_table)
no_Avastin_patients <- sum(Avastin_data_table[,c(3)], na.rm = T)
```


</div>
</div>



<br>


Scatterplot for Laser again showing no real correlation


<div class = "row">
<div class="column-left">
```{r, warning=FALSE}
Laser_data <- metafor_all_data %>%
filter(grepl("Laser", `trt`)) 

Laser_data_mod <- 
            metafor::escalc(data = Laser_data,  
                            measure = "PFT", xi=xi, ni=ni, 
                            slab = paste(authors, trt), add = 0)

Laser_modplot <- metafor::rma(yi, vi, 
                            mods = ~ g.age,
                            data = Laser_data_mod,                                          
                            method = "REML")

Laser_predict <- 
          predict(Laser_modplot, newmods=c(20:35),     
          transf=transf.ipft.hm, 
          targ=list(ni=Laser_data_mod$ni))


wi=1/sqrt(Laser_data_mod$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(Laser_data_mod$g.age,transf.ipft.hm(Laser_data_mod$yi,targ=list(ni=Laser_data_mod$ni)),cex=size,pch=1,las=1,xlab="Gestational age (wks)", ylab="Proportional effect size")
lines(20:35,Laser_predict$pred,col="navy")
lines(20:35,Laser_predict$ci.lb,lty="dashed",col="maroon")
lines(20:35,Laser_predict$ci.ub,lty="dashed",col="maroon")
ids=c(1:18)
text(Laser_data_mod$g.age[ids],transf.ipft.hm(Laser_data_mod$yi,targ=list(ni=Laser_data_mod$ni))[ids],ids,cex=0.8,pos=1)
```

</div>
<div class="column-right">

```{r}
Laser_data_table <- Laser_data_mod[c(3,6,7,8)]
Laser_data_table <- mutate(Laser_data_table, 
                        study = 1:nrow(Laser_data_table))
Laser_data_table <- Laser_data_table[,c(5,1,2,3,4)]
colnames(Laser_data_table) <- c("Study", "Authors", "Sample size", "Gestational age", "Birth weight")

Laser_study_table <- 
  Laser_data_table %>%
  kbl(caption = "Laser") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=9)

Laser_study_table

no_Laser_studies <- nrow(Laser_data_table)
no_Laser_patients <- sum(Laser_data_table[,c(3)], na.rm = T)
```


</div>
</div>


```{r eval=FALSE, include=FALSE}
### run permutation test to validate the robustness of the meta-regression models
metafor::permutest(Lucentis_modplot)
metafor::permutest(Avastin_modplot)
metafor::permutest(Laser_modplot)
```


<br>


Recall this is a conventional frequentist meta-analysis/meta-regression and so does not compare relative effect sizes of treatments. It is simply to show whether gestational age provides a significant contributing factor towards the positive response effect size for each treatment individually. Bayesian analsysis is needed to compare one treatement vs another - coming next.

<br>

I have performed a permutation test for the above conventional frequentist meta-regression to validate the robustness of the models.  

 
<br><br>




Compare GA group means between Avastin and Lucentis (with normality and equal variance test) using student t-test


```{r}
################ Stats tests ################################
#############################################################
### test normality of data first with shapiro-wilk
shapiro.test(df$`mean gestational age - Lucentis (wks)`)
### F-test to check for equal variances (condition of using conventional t-test)
var.test(df$`mean gestational age - Avastin (wks)`, df$`mean gestational age - Lucentis (wks)`)

### above all ok so perform student t-test to compare group means
means <- t.test(df$`mean gestational age - Avastin (wks)`, df$`mean gestational age - Lucentis (wks)`, var.equal = T)
means
```




<br><br><br>


## Bayesian network meta-analysis
<br>

### Summary of network
<br>

```{r message=FALSE, warning=FALSE, include=FALSE}
df_long2 <- df_long[c(2,4,5,6,7,8)]

df_long2 <- na.omit(df_long2)

colnames(df_long2)[2] <- "treatment"

### for relative risk better to use non-responders rather than responders
df_long2 <- mutate(df_long2, responders=sampleSize-responders)

network.mr <- mtc.network(df_long2, 
                          studies = df_long2[c(1,2,5,6)],
                          description = "Bayesian NMA")
```

```{r}
summary(network.mr)
#summary(anohe)
```

<br>
<br>

Network plot (line thickness connecting nodes denotes the number studies for each comparison)
```{r}
plot(network.mr)
```


<br>

## Meta-Regression

<br><br>

### Gestational age as a regressor in the Bayesian meta-regression (potential moderator)

<br>

```{r}

classes <- list("Laser"=c("Laser"),
                "Avastin"=c("Avastin"),
                "Lucentis"=c("Lucentis"),
                "Eylea"=c("Eylea"))


set.seed(42)

model.mr <- mtc.model(network.mr, 
                      likelihood = "binom", 
                      link = "log", 
                      linearModel = "random", 
                      n.chain = 4, 
                      type = "regression", 
                      regressor = list(coefficient="unrelated",variable="gest.age", control="Laser"))

#################################################################
# By default, the model will have 4 chains - generate a seed for each - see gemtc documentation
seeds <- sample.int(4, n = .Machine$integer.max)
# Apply JAGS RNG settings to each chain
model.mr$inits <- mapply(c, model.mr$inits, list(
list(.RNG.name="base::Wichmann-Hill", .RNG.seed=seeds[1]),
list(.RNG.name="base::Marsaglia-Multicarry", .RNG.seed=seeds[2]),
list(.RNG.name="base::Super-Duper", .RNG.seed=seeds[3]),
list(.RNG.name="base::Mersenne-Twister", .RNG.seed=seeds[4])), SIMPLIFY=FALSE)
#################################################################

mcmc.mr <- mtc.run(model.mr, 
                n.adapt = 8000, 
                n.iter = 200000, 
                thin = 10)
summary(mcmc.mr)
```

```{r,fig.width=8, fig.height=12,fig.align="center", include=FALSE}
plot(mcmc.mr)
```

```{r eval=FALSE, fig.height=6, fig.width=6, include=FALSE}
gelman.plot(mcmc.mr)
gelman.diag(mcmc.mr, multivariate = F)
PSRF <- gelman.diag(mcmc.mr)$mpsrf
print(paste0("PSRF = ", round(PSRF, digits = 4)))
```

```{r eval=FALSE, include=FALSE}
nodesplit <- mtc.nodesplit(network.mr,     
                           linearModel = "random",
                           n.adapt = 8000,
                           n.iter = 200000,
                           thin = 10)
```

```{r eval=FALSE, include=FALSE}
summary(nodesplit)
names(nodesplit)
```

```{r,fig.align="center"}
ranks.mr <- rank.probability(mcmc.mr, preferredDirection = -1)
rankquant.mr <- rank.quantiles(ranks.mr, probs=c("2.5%"=0.025, "50%"=0.5, "97.5%"=0.975))
```

```{r eval=FALSE, fig.align="center", fig.cap="A", include=FALSE}
plot(ranks.mr, beside=TRUE) # plot a 'rankogram'
```

```{r eval=FALSE, include=FALSE , fig.align="center"}
print(gemtc::sucra(ranks.mr), digits=3)
```



<br><br>

Here are comparative forest plots comparing treatment effects against one another. I have fixed the gestational age to 26 weeks for illustrative purposes but the range of gestational ages is shown in the covariate plots below. 

<br>

```{r, fig.width=8, fig.height=14,fig.align="center",fig.asp=0.30}

weeks <- 28.9
gemtc::forest(relative.effect(mcmc.mr, t1 = "Laser", covariate = weeks))
gemtc::forest(relative.effect(mcmc.mr, t1 = "Lucentis", covariate = weeks))
gemtc::forest(relative.effect(mcmc.mr, t1 = "Eylea", covariate = weeks))
gemtc::forest(relative.effect(mcmc.mr, t1 = "Avastin", covariate = weeks))
```


<br>

## Covariate plots for gestational age

<br>

Covariate plots below are in log(RR) units.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr, t1="Avastin", t2="Lucentis", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```




<br>

From the above covariate plot we can see there are no statistically significant differences between Avastin and Lucentis outcomes across gestational ages (credible intervals always incorporate zero).


There is a downward slope suggesting Lucentis may be associated with greater risk of requiring re-treatment at lower GA (on the threshold for significance but not quite significant) but then improves to be associated with lower risk at higher GA, although none of these associations reach statistical significance unfortunately.



<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr, t1="Laser", t2="Lucentis", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>
Comparison of Lucentis versus Laser shows some statistical difference, with Lucentis associated with a higher relative risk of needing re-treatment for a small range of gestational ages between 25.8 weeks and 28.9 weeks GA. At all other GA there is no statistically significant differences between the two treatment modalities.


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr, t1="Laser", t2="Avastin", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>

Interestingly here we are seeing quite a difference between Avastin and Laser at higher GA's with Avastin being associated with a statististically significant increased risk of requiring re-treatment compared to Laser (over 26.4 weeks) and steadily increases in risk with increasing GA.


At low GA's (<=26.4 wks) there are no statistically significant differences between Avastin and Laser.

<br>




```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr, t1="Laser", t2="Eylea", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>

All Eylea combination plots look the same as above whereby the are no statistically significant differences between Eylea and all other treatments for any GA. (N.B only Eylea vs Laser shown here but all other Eylea comparisons look very similar). 




<br><br>


***
Summary


For lower GA there are no statistically significant differences in outcomes between any treatment modality. There are indications Avastin may  
be associated with a lower risk of requiring re-treatment at low GA compared to Lucentis and Laser however the results do not quite reach statistical significance.


Between 25.8 weeks and 28.9 weeks GA, Lucentis is associated with a statistically significant increased risk of requiring re-treatment compared to Laser. 


At GA greater than 26.4 weeks, Avastin is associated with a statistically significant increased risk of requiring re-treatment compared to Laser.



There are no statistically significant results for Eylea comparisons. 

***


<br><br>











<br>

### Birth weight as a regressor in the Bayesian meta-regression (potential moderator)

<br>

```{r}

classes <- list("Laser"=c("Laser"),
                "Avastin"=c("Avastin"),
                "Lucentis"=c("Lucentis"),
                "Eylea"=c("Eylea"))


set.seed(42)

model.mr2 <- mtc.model(network.mr, 
                      likelihood = "binom", 
                      link = "log", 
                      linearModel = "random", 
                      n.chain = 4, 
                      type = "regression", 
                      regressor = list(coefficient="unrelated",variable="b.weight", control="Laser"))

#################################################################
# By default, the model will have 4 chains - generate a seed for each - see gemtc documentation
seeds <- sample.int(4, n = .Machine$integer.max)
# Apply JAGS RNG settings to each chain
model.mr2$inits <- mapply(c, model.mr2$inits, list(
list(.RNG.name="base::Wichmann-Hill", .RNG.seed=seeds[1]),
list(.RNG.name="base::Marsaglia-Multicarry", .RNG.seed=seeds[2]),
list(.RNG.name="base::Super-Duper", .RNG.seed=seeds[3]),
list(.RNG.name="base::Mersenne-Twister", .RNG.seed=seeds[4])), SIMPLIFY=FALSE)
#################################################################

mcmc.mr2 <- mtc.run(model.mr2, 
                n.adapt = 8000, 
                n.iter = 200000, 
                thin = 10)
summary(mcmc.mr2)
```

```{r,fig.width=8, fig.height=12,fig.align="center", include=FALSE}
plot(mcmc.mr2)
```

```{r, fig.height=6, fig.width=6, eval=FALSE, include=FALSE}
gelman.plot(mcmc.mr2)
gelman.diag(mcmc.mr2, multivariate = F)
PSRF <- gelman.diag(mcmc.mr2)$mpsrf
print(paste0("PSRF = ", round(PSRF, digits = 4)))
```

```{r,fig.align="center"}
ranks.mr2 <- rank.probability(mcmc.mr2, preferredDirection = -1)
rankquant.mr2 <- rank.quantiles(ranks.mr2, probs=c("2.5%"=0.025, "50%"=0.5, "97.5%"=0.975))
```

```{r eval=FALSE, fig.align="center", fig.cap="A", include=FALSE}
plot(ranks.mr2, beside=TRUE) # plot a 'rankogram'
```

```{r eval=FALSE, include=FALSE , fig.align="center"}
print(gemtc::sucra(ranks.mr2), digits=3)
```



<br><br>

Here are comparative forest plots comparing treatment effects against one another. I have fixed the birth weight to 1000g for illustrative purposes. 

<br>

```{r, fig.width=8, fig.height=14,fig.align="center",fig.asp=0.30}

weight <- 1400
gemtc::forest(relative.effect(mcmc.mr2, t1 = "Laser", covariate = weight))
gemtc::forest(relative.effect(mcmc.mr2, t1 = "Lucentis", covariate = weight))
gemtc::forest(relative.effect(mcmc.mr2, t1 = "Eylea", covariate = weight))
gemtc::forest(relative.effect(mcmc.mr2, t1 = "Avastin", covariate = weight))
```


<br>

## Covariate plots for birth weight

<br>

Covariate plots below are in log(RR) units.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr2, t1="Avastin", t2="Lucentis", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

There are no significant differences between re-treatment outcomes between Avastin and Lucentis across various birth weights.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr2, t1="Laser", t2="Lucentis", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

There are no statistically significant differences in re-treatment outcomes between Lucentis and Laser except for a middle range of birth weights (841g-1243g). This narrow range just meets the criteria for statistical significance.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr2, t1="Laser", t2="Avastin", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

At lower birth weights there are no differences in re-treatment outcomes between Avastin and Laser. Above 911g Avastin is associated with a statistically significant increased risk of requiring re-treatment compared to Laser. 


For example at a borth weight of 1400g, treatment with Avastin is associated with 8.6 times (CrI=1.3, 75) increased risk of requiring re-treatment than Laser. Note, the credible intervals of 1.3-75 means this is only just significant with fairly wide intervals. Also note covariate plots are in natural log units. 

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr2, t1="Laser", t2="Eylea", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

Again, there are no statistically significant differences in re-treatment outcomes evident between Eylea and any other treatment modality.

<br>





