---
title: "ROP meta-analysis"
author: "Aman Josan"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document: 
    self_contained: no
    
    

editor_options: 
  chunk_output_type: inline
---




<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-right{
  float: right;
  width: 50%;
  text-align: right;
}
</style>





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, comment=NA, cache.rebuild = TRUE)
library(dplyr)
library(readxl)
library(robumeta)
library(gemtc)
library(rjags)
library(reshape2)
library(kableExtra)
library(metafor)
library(dmetar)
library(tidyverse)
library(here)
```


<br><br>


### Type 1 ROP - Gestational age, birth weight and postmenstrual age as potential moderators
<br>

```{r, message=FALSE, warning=FALSE}
data <- suppressMessages(read_excel(here("./data/ROP data 29122022-V9.xlsx"),sheet = 1, range = "b3:ba33", col_names = T))

df <- data %>% select("study",
                      "treatment-Avastin",
                      "treatment-Lucentis",
                      "treatment-Eylea",
                      "treatment-laser",
                      "study design", 
                      "no. Eyes-Avastin",
                      "no. Eyes-Lucentis",
                      "no. Eyes-Eylea",
                      "no. Eyes-laser",
                      "no. retreated-Avastin",
                      "no. retreated-Lucentis",
                      "no. retreated-Eylea",
                      "no. retreated-laser",
                      "mean gestational age - Avastin (wks)",
                      "mean gestational age - Lucentis (wks)",
                      "mean gestational age - Eylea (wks)",
                      "mean gestational age - Laser (wks)",
                      "mean birth weight - Avastin (wks)",
                      "mean birth weight - Lucentis (wks)",
                      "mean birth weight - Eylea (wks)",
                      "mean birth weight - Laser (wks)",
                      "mean PMA at tx (weeks)- Avastin",
                      "mean PMA at tx (weeks)- Lucentis",
                      "mean PMA at tx (weeks)- Eylea",
                      "mean PMA at tx (weeks)- Laser"
                      ) 

df[df == "n/a" ] <- NA

df[,c(7:26)] <- suppressWarnings(lapply(df[,c(7:26)], as.numeric))



df$`n_responders-Avastin` <- df$`no. Eyes-Avastin` - df$`no. retreated-Avastin`
df$`n_responders-Lucentis` <- df$`no. Eyes-Lucentis` - df$`no. retreated-Lucentis`
df$`n_responders-Eylea` <- df$`no. Eyes-Eylea` - df$`no. retreated-Eylea`
df$`n_responders-laser` <- df$`no. Eyes-laser` - df$`no. retreated-laser`

df <- df[,-c(11:14)]

#### change from trade names to drug names 
#Avastin - Bevacizumab
#Lucentis - Ranibizumab
#Eylea - Aflibercept
#Laser
names(df)[names(df) == 'treatment-Avastin'] <- 'treatment-Bevacizumab'
names(df)[names(df) == 'treatment-Lucentis'] <- 'treatment-Ranibizumab'
names(df)[names(df) == 'treatment-Eylea'] <- 'treatment-Aflibercept'
names(df)[names(df) == 'treatment-laser'] <- 'treatment-Laser'
names(df)[names(df) == 'no. Eyes-Avastin'] <- 'no. Eyes-Bevacizumab'
names(df)[names(df) == 'no. Eyes-Lucentis'] <- 'no. Eyes-Ranibizumab'
names(df)[names(df) == 'no. Eyes-Eylea'] <- 'no. Eyes-Aflibercept'
names(df)[names(df) == 'no. Eyes-laser'] <- 'no. Eyes-Laser'
names(df)[names(df) == 'n_responders-Avastin'] <- 'no. responders-Bevacizumab'
names(df)[names(df) == 'n_responders-Lucentis'] <- 'no. responders-Ranibizumab'
names(df)[names(df) == 'n_responders-Eylea'] <- 'no. responders-Aflibercept'
names(df)[names(df) == 'n_responders-laser'] <- 'no. responders-Laser'



df <- mutate(df, study_id = 1:nrow(df)) # This adds a study id
df <- df[c(27,1:26)] # rearranges columns

df_table <- df[,-c(1)]
### print table of studies
df_table %>%
  kbl(caption = "All studies which considered type 1 ROP") %>%
  kable_classic(full_width = T, html_font = "Cambria", font_size=9) %>%
  kable_styling("striped") 

no_studies <- nrow(df_table)
no_patients <- sum(df[,c(8:11)], na.rm = T)
```

<br><br>

number of studies=`r no_studies`


Number of patients=`r no_patients`

<br><br>


## Event count data and conventional meta-analysis of proportions


<br>


### Transform data (Dichotomous data)

<br>

Here we define the number of responders as the number of patients treated minus the number of patients requiring retreatment. Hence the measure of treatment effectiveness we are applying is the number of patients requiring no retreatment.


Since we are dealing with raw prevalence (or event rate) data (i.e. sample size and number of treatment responders) rather than correlations (i.e. Pearsons r for treatment correlations) we need to perform a meta-analysis of proportions.  Most meta-analysis in the literature deal with correlations and p-values.



The proportion of treatment responders ($p$) is given by

$$\begin{align}
p = \frac{x}{n}
\end{align}$$

<br>

where $x$ is the number of treatment responders in a study and $n$ is the total number treated with that treatment in the study.



Again data is quite skewed (as is normally the case in number of event data), hence use double arcsine transformation for the conventional meta-analysis again.




```{r, results="hide"}
### prep dataframe for analysis by having t,n,r columns
df2 <- df

df2$t.1 <- 1 #"Avastin - Bevacizumab" 
df2$t.2 <- 2 #"Lucentis - Ranibizumab" 
df2$t.3 <- 3 #"Eylea - Aflibercept"
df2$t.4 <- 4 #"Laser"

df2 <- df2[c(1,2,28:31,8:11,24:27,12:15,16:19,20:23)]

colnames(df2)[7] <- "n.1"
colnames(df2)[8] <- "n.2"
colnames(df2)[9] <- "n.3"
colnames(df2)[10] <- "n.4"
colnames(df2)[11] <- "r.1"
colnames(df2)[12] <- "r.2"
colnames(df2)[13] <- "r.3"
colnames(df2)[14] <- "r.4"
colnames(df2)[15] <- "ga.1"
colnames(df2)[16] <- "ga.2"
colnames(df2)[17] <- "ga.3"
colnames(df2)[18] <- "ga.4"
colnames(df2)[19] <- "bw.1"
colnames(df2)[20] <- "bw.2"
colnames(df2)[21] <- "bw.3"
colnames(df2)[22] <- "bw.4"
colnames(df2)[23] <- "pma.1"
colnames(df2)[24] <- "pma.2"
colnames(df2)[25] <- "pma.3"
colnames(df2)[26] <- "pma.4"


studyID <- df2[,c(1,2)]
df2 <- suppressWarnings(data.frame(lapply(df2,as.numeric)))
df2 <- df2[,-c(2)]
```

```{r, include=FALSE}
### reshape to group by treatment type
df_long <- reshape(df2, 
  varying = c("t.1","t.2","t.3","t.4",
              "n.1","n.2","n.3","n.4",
              "r.1","r.2","r.3","r.4",
              "ga.1","ga.2","ga.3","ga.4",
              "bw.1","bw.2","bw.3","bw.4",
              "pma.1","pma.2","pma.3","pma.4"), 
  timevar = "treatment",
  direction = "long")

df_long <- df_long[,c(1,3,4,5,6,7,8)] %>% arrange(study_id)

df_long <- df_long[!is.na(df_long$r), ]
df_long <- df_long[!is.na(df_long$n), ]

colnames(df_long) <- c("study_id", "treatment", "sampleSize", "responders", "gest.age", "b.weight", "pma") 

df_long <- merge(x = df_long, y = studyID, by = "study_id", all.x = TRUE)
df_long <- df_long[c(1,8,2,3,4,5,6,7)]
df_long$treatment <- as.character(df_long$treatment)
df_long$trt <- ""
df_long$trt[df_long$treatment=="1"] <- "Bevacizumab"
df_long$trt[df_long$treatment=="2"] <- "Ranibizumab"
df_long$trt[df_long$treatment=="3"] <- "Aflibercept"
df_long$trt[df_long$treatment=="4"] <- "Laser"
df_long <- df_long[c(1,2,3,9,4,5,6,7,8)]
```

```{r, include=FALSE, warning=FALSE}
### use metafor package to plot forest, funnel plots and - not available in gemtc
metafor_data <- df_long
colnames(metafor_data) <- c("study", "authors","treatment","trt","ni","xi","g.age","b.weight","pma")
metafor_data <- mutate(metafor_data, id = 1:nrow(metafor_data))
metafor_all_data <- metafor_data[c(10,1,2,4,5,6,7,8,9)]



### use escalc to find effect size in terms of freeman & Tukey proportions
metafor_all_df <- metafor::escalc(data=metafor_all_data, measure="PFT", xi=xi, ni=ni, slab=paste(authors, trt), add = 0)


### perform mixed effects model on each group
metafor_all_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight+pma,data=metafor_all_df, method = "REML")

metafor_Avastin_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight+pma,data=metafor_all_df, method = "REML", subset=(trt=="Bevacizumab"))

metafor_Lucentis_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight+pma,data=metafor_all_df, method = "REML", subset=(trt=="Ranibizumab"))

metafor_Laser_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight+pma,data=metafor_all_df, method = "REML", subset=(trt=="Laser"))

metafor_Eylea_res_mod <- 
  metafor::rma(yi,vi,mods = ~ g.age+b.weight,data=metafor_all_df, method = "REML", subset=(trt=="Aflibercept"))

metafor_Eylea_res_mod2 <- 
  metafor::rma(yi,vi,mods = ~ pma,data=metafor_all_df, method = "REML", subset=(trt=="Aflibercept"))

```



```{r}
Avastin_stats <- 
  data.frame(capture.output(metafor_Avastin_res_mod)) %>%
  kbl(caption = "Avastin") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Avastin_stats


```


```{r}
Lucentis_stats <- data.frame(capture.output(metafor_Lucentis_res_mod)) %>%
  kbl(caption = "Lucentis") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Lucentis_stats

```


```{r}
Laser_stats <- 
  data.frame(capture.output(metafor_Laser_res_mod)) %>%
  kbl(caption = "Laser") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Laser_stats
```


```{r}
Eylea_stats <- data.frame(capture.output(metafor_Eylea_res_mod)) %>%
  kbl(caption = "Eylea") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Eylea_stats
```


```{r}
Eylea_stats2 <- data.frame(capture.output(metafor_Eylea_res_mod2)) %>%
  kbl(caption = "Eylea") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=12)
Eylea_stats2
```





<br><br>

Looking at the lines titled "test of moderators" and "g.age", "b.weight" and  "pma" under "model results" we see that the model predictions find neither gestational age, birth weight nor mean postmenstrual age at treatment have a significant moderating factor. 


For example for those treated with laser, the p-value for the combined moderating effect of GA, BW & pma has a p-value=0.7256 and with individual moderating effect p-values for GA=0.8067 and BW=0.8489 and pma=0.2786.



<br>

Hence all treatments now have non-significant correlations for gestational age and birth weight and PMA at treatment as moderators for effect size. Eylea moderator analysis had to be split into two parts.The first analyses the effects of all combinations of g.age and birth weight individually and together on the effect size. The second analyses the moderating effect of PMA alone on the effect size. It was not possible to combine PMA with g.age and birth weight to analyse combinations of all moderators as there are too few PMA samples for Eylea. No moderating effects were found for Eylea. 

<br>


Performing a frequentist meta-regression and generating a scatter plot with regression line and confidence intervals to see the relationship although now there is no significant correlation so these plots can be mostly ignored. Here I've only shown plots for gestational age as birth weight plots look pretty similar. I can plot them if needed but as they are non-significant I haven't bothered here.


<br><br>

Hence, for Lucentis, flat-ish line shows there is no real correlation to GA (especially when taking confidence interval widths into account, i.e. line could be completely flat):

<div class = "row">
<div class="column-left">
```{r, warning=FALSE}
Lucentis_data <- metafor_all_data %>%
filter(grepl("Ranibizumab", `trt`)) 

Lucentis_data_mod <- 
            metafor::escalc(data = Lucentis_data,  
                            measure = "PFT", xi=xi, ni=ni, 
                            slab = paste(authors, trt), add = 0)

Lucentis_modplot <- metafor::rma(yi, vi, 
                            mods = ~ g.age,
                            data = Lucentis_data_mod,                                         
                            method = "REML")

Lucentis_predict <- 
          predict(Lucentis_modplot, newmods=c(20:35),     
          transf=transf.ipft.hm, 
          targ=list(ni=Lucentis_data_mod$ni))


wi=1/sqrt(Lucentis_data_mod$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(Lucentis_data_mod$g.age,transf.ipft.hm(Lucentis_data_mod$yi,targ=list(ni=Lucentis_data_mod$ni)),cex=size,pch=1,las=1,xlab="Gestational age (wks)", ylab="Proportional effect size")
lines(20:35,Lucentis_predict$pred,col="navy")
lines(20:35,Lucentis_predict$ci.lb,lty="dashed",col="maroon")
lines(20:35,Lucentis_predict$ci.ub,lty="dashed",col="maroon")
ids=c(1:18)
text(Lucentis_data_mod$g.age[ids],transf.ipft.hm(Lucentis_data_mod$yi,targ=list(ni=Lucentis_data_mod$ni))[ids],ids,cex=0.8,pos=1)
```

</div>
<div class="column-right">

```{r}
Lucentis_data_table <- Lucentis_data_mod[c(3,6,7,8,9)]
Lucentis_data_table <- mutate(Lucentis_data_table, 
                        study = 1:nrow(Lucentis_data_table))
Lucentis_data_table <- Lucentis_data_table[,c(6,1,2,3,4,5)]
colnames(Lucentis_data_table) <- c("Study", "Authors", "Sample size", "Gestational age (weeks)","Birth weight (g)","PMA at treatment (weeks")

Luc_study_table <- 
  Lucentis_data_table %>%
  kbl(caption = "Ranibizumab studies and assigned ID's") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=9)

Luc_study_table

no_Luc_studies <- nrow(Lucentis_data_table)
no_Luc_patients <- sum(Lucentis_data_table[,c(3)], na.rm = T)
```
</div>
</div>


<br>

N.B. Larger circles represent larger studies. 


<br><br>


Scatterplot for Avastin again showing a flat-ish line representing no real correlation (completely flat line cannot be excluded when looking at the CI)


<div class = "row">
<div class="column-left">
```{r, warning=FALSE}
Avastin_data <- metafor_all_data %>%
filter(grepl("Bevacizumab", `trt`)) 

Avastin_data_mod <- 
            metafor::escalc(data = Avastin_data,  
                            measure = "PFT", xi=xi, ni=ni, 
                            slab = paste(authors, trt), add = 0)

Avastin_modplot <- metafor::rma(yi, vi, 
                            mods = ~ g.age,
                            data = Avastin_data_mod,                                          
                            method = "REML")

Avastin_predict <- 
          predict(Avastin_modplot, newmods=c(20:35),     
          transf=transf.ipft.hm, 
          targ=list(ni=Avastin_data_mod$ni))


wi=1/sqrt(Avastin_data_mod$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(Avastin_data_mod$g.age,transf.ipft.hm(Avastin_data_mod$yi,targ=list(ni=Avastin_data_mod$ni)),cex=size,pch=1,las=1,xlab="Gestational age (wks)", ylab="Proportional effect size")
lines(20:35,Avastin_predict$pred,col="navy")
lines(20:35,Avastin_predict$ci.lb,lty="dashed",col="maroon")
lines(20:35,Avastin_predict$ci.ub,lty="dashed",col="maroon")
ids=c(1:24)
text(Avastin_data_mod$g.age[ids],transf.ipft.hm(Avastin_data_mod$yi,targ=list(ni=Avastin_data_mod$ni))[ids],ids,cex=0.8,pos=1)
```

</div>
<div class="column-right">

```{r}
Avastin_data_table <- Avastin_data_mod[c(3,6,7,8,9)]
Avastin_data_table <- mutate(Avastin_data_table, 
                        study = 1:nrow(Avastin_data_table))
Avastin_data_table <- Avastin_data_table[,c(6,1,2,3,4,5)]
colnames(Avastin_data_table) <- c("Study", "Authors", "Sample size", "Gestational age (weeks)", "Birth weight (g)", "PMA at treatment (weeks)")

Avastin_study_table <- 
  Avastin_data_table %>%
  kbl(caption = "Bevacizumab") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=9)

Avastin_study_table

no_Avastin_studies <- nrow(Avastin_data_table)
no_Avastin_patients <- sum(Avastin_data_table[,c(3)], na.rm = T)
```


</div>
</div>



<br>


Scatterplot for Laser again showing no real correlation (completely flat line cannot be excluded when looking at the CI)


<div class = "row">
<div class="column-left">
```{r, warning=FALSE}
Laser_data <- metafor_all_data %>%
filter(grepl("Laser", `trt`)) 

Laser_data_mod <- 
            metafor::escalc(data = Laser_data,  
                            measure = "PFT", xi=xi, ni=ni, 
                            slab = paste(authors, trt), add = 0)

Laser_modplot <- metafor::rma(yi, vi, 
                            mods = ~ g.age,
                            data = Laser_data_mod,                                          
                            method = "REML")

Laser_predict <- 
          predict(Laser_modplot, newmods=c(20:35),     
          transf=transf.ipft.hm, 
          targ=list(ni=Laser_data_mod$ni))


wi=1/sqrt(Laser_data_mod$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(Laser_data_mod$g.age,transf.ipft.hm(Laser_data_mod$yi,targ=list(ni=Laser_data_mod$ni)),cex=size,pch=1,las=1,xlab="Gestational age (wks)", ylab="Proportional effect size")
lines(20:35,Laser_predict$pred,col="navy")
lines(20:35,Laser_predict$ci.lb,lty="dashed",col="maroon")
lines(20:35,Laser_predict$ci.ub,lty="dashed",col="maroon")
ids=c(1:27)
text(Laser_data_mod$g.age[ids],transf.ipft.hm(Laser_data_mod$yi,targ=list(ni=Laser_data_mod$ni))[ids],ids,cex=0.8,pos=1)
```

</div>
<div class="column-right">

```{r}
Laser_data_table <- Laser_data_mod[c(3,6,7,8,9)]
Laser_data_table <- mutate(Laser_data_table, 
                        study = 1:nrow(Laser_data_table))
Laser_data_table <- Laser_data_table[,c(6,1,2,3,4,5)]
colnames(Laser_data_table) <- c("Study", "Authors", "Sample size", "Gestational age (weeks)", "Birth weight (g)", "PMA at treatment (weeks)")

Laser_study_table <- 
  Laser_data_table %>%
  kbl(caption = "Laser") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=9)

Laser_study_table

no_Laser_studies <- nrow(Laser_data_table)
no_Laser_patients <- sum(Laser_data_table[,c(3)], na.rm = T)
```


</div>
</div>



<br>

Scatterplot for Eylea again showing no real correlation (completely flat line cannot be excluded when looking at the CI)


<div class = "row">
<div class="column-left">
```{r, warning=FALSE}
Eylea_data <- metafor_all_data %>%
filter(grepl("Aflibercept", `trt`)) 

Eylea_data_mod <- 
            metafor::escalc(data = Eylea_data,  
                            measure = "PFT", xi=xi, ni=ni, 
                            slab = paste(authors, trt), add = 0)

Eylea_modplot <- metafor::rma(yi, vi, 
                            mods = ~ g.age,
                            data = Eylea_data_mod,                                          
                            method = "REML")

Eylea_predict <- 
          predict(Eylea_modplot, newmods=c(20:35),     
          transf=transf.ipft.hm, 
          targ=list(ni=Eylea_data_mod$ni))


wi=1/sqrt(Eylea_data_mod$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(Eylea_data_mod$g.age,transf.ipft.hm(Eylea_data_mod$yi,targ=list(ni=Eylea_data_mod$ni)),cex=size,pch=1,las=1,xlab="Gestational age (wks)", ylab="Proportional effect size")
lines(20:35,Eylea_predict$pred,col="navy")
lines(20:35,Eylea_predict$ci.lb,lty="dashed",col="maroon")
lines(20:35,Eylea_predict$ci.ub,lty="dashed",col="maroon")
ids=c(1:27)
text(Eylea_data_mod$g.age[ids],transf.ipft.hm(Eylea_data_mod$yi,targ=list(ni=Eylea_data_mod$ni))[ids],ids,cex=0.8,pos=1)
```

</div>
<div class="column-right">

```{r}
Eylea_data_table <- Eylea_data_mod[c(3,6,7,8,9)]
Eylea_data_table <- mutate(Eylea_data_table, 
                        study = 1:nrow(Eylea_data_table))
Eylea_data_table <- Eylea_data_table[,c(6,1,2,3,4,5)]
colnames(Eylea_data_table) <- c("Study", "Authors", "Sample size", "Gestational age (weeks)", "Birth weight (g)", "PMA at treatment (weeks)")

Eylea_study_table <- 
  Eylea_data_table %>%
  kbl(caption = "Aflibercept") %>%
  kable_classic(full_width = F, html_font = "Cambria", 
                font_size=9)

Eylea_study_table

no_Eylea_studies <- nrow(Eylea_data_table)
no_Eylea_patients <- sum(Eylea_data_table[,c(3)], na.rm = T)
```


</div>
</div>


```{r eval=FALSE}
### run permutation test to validate the robustness of the meta-regression models
metafor::permutest(Lucentis_modplot)
metafor::permutest(Avastin_modplot)
metafor::permutest(Laser_modplot)
metafor::permutest(Eylea_modplot)
```


<br>


Recall this is a conventional frequentist meta-analysis/meta-regression and so does not compare relative effect sizes of treatments. It is simply to show whether gestational age, birth weight and/or postmenstrual age provides a significant contributing factor towards the positive response effect size for each treatment individually. Bayesian analsysis is needed to compare one treatement vs another - coming next.

<br>

I have performed a permutation test for the above conventional frequentist meta-regression to validate the robustness of the models across all moderators.  

 
<br><br>




Compare GA group means between Avastin and Lucentis (with normality and equal variance test) using student t-test


```{r}
################ Stats tests ################################
#############################################################
### test normality of data first with shapiro-wilk
shapiro.test(df$`mean gestational age - Lucentis (wks)`)
### F-test to check for equal variances (condition of using conventional t-test)
var.test(df$`mean gestational age - Avastin (wks)`, df$`mean gestational age - Lucentis (wks)`)

### above all ok so perform student t-test to compare group means
means <- t.test(df$`mean gestational age - Avastin (wks)`, df$`mean gestational age - Lucentis (wks)`, var.equal = T)
means
```




<br><br><br>


## Bayesian network meta-analysis
<br><br>

## Meta-Regression



```{r message=FALSE, warning=FALSE, include=FALSE}
df_long2 <- df_long[c(2,4,5,6,7,8,9)]

colnames(df_long2)[2] <- "treatment"

### for relative risk better to use non-responders rather than responders
df_long2 <- mutate(df_long2, responders=sampleSize-responders)
```













<br><br>

### Gestational age as a regressor in the Bayesian meta-regression (potential moderator)

<br>


```{r}
df_long2_g.age <- na.omit(df_long2[,c(1:5)])

network.mr_g.age <- mtc.network(df_long2_g.age, 
                          studies = df_long2_g.age[c(1,2,5)],
                          description = "Bayesian NMA")
```

```{r}
summary(network.mr_g.age)
#summary(anohe)
```

<br>
<br>

Network plot (line thickness connecting nodes denotes the number studies for each comparison)
```{r}
plot(network.mr_g.age)
```


<br>

```{r}
set.seed(42)


model.mr_g.age <- mtc.model(network.mr_g.age, 
                      likelihood = "binom", 
                      link = "log", 
                      linearModel = "random", 
                      n.chain = 4, 
                      type = "regression", 
                      regressor = list(coefficient="unrelated",variable="gest.age", control="Laser"))

#################################################################
# By default, the model will have 4 chains - generate a seed for each - see gemtc documentation
seeds <- sample.int(4, n = .Machine$integer.max)
# Apply JAGS RNG settings to each chain
model.mr_g.age$inits <- mapply(c, model.mr_g.age$inits, list(
list(.RNG.name="base::Wichmann-Hill", .RNG.seed=seeds[1]),
list(.RNG.name="base::Marsaglia-Multicarry", .RNG.seed=seeds[2]),
list(.RNG.name="base::Super-Duper", .RNG.seed=seeds[3]),
list(.RNG.name="base::Mersenne-Twister", .RNG.seed=seeds[4])), SIMPLIFY=FALSE)
#################################################################

mcmc.mr_g.age <- mtc.run(model.mr_g.age, 
                n.adapt = 8000, 
                n.iter = 200000, 
                thin = 10)
summary(mcmc.mr_g.age)
```

```{r,fig.width=8, fig.height=12,fig.align="center"}
plot(mcmc.mr_g.age)
```

```{r eval=FALSE, fig.height=6, fig.width=6}
gelman.plot(mcmc.mr_g.age)
gelman.diag(mcmc.mr_g.age, multivariate = F)
PSRF <- gelman.diag(mcmc.mr_g.age)$mpsrf
print(paste0("PSRF = ", round(PSRF, digits = 4)))
```

```{r eval=FALSE}
nodesplit <- mtc.nodesplit(network.mr_g.age,     
                           linearModel = "random",
                           n.adapt = 8000,
                           n.iter = 200000,
                           thin = 10)
```

```{r eval=FALSE}
summary(nodesplit)
names(nodesplit)
```

```{r,fig.align="center"}
ranks.mr_g.age <- rank.probability(mcmc.mr_g.age, preferredDirection = -1)
rankquant.mr_g.age <- rank.quantiles(ranks.mr_g.age, probs=c("2.5%"=0.025, "50%"=0.5, "97.5%"=0.975))
```

```{r eval=FALSE, fig.align="center", fig.cap="A"}
plot(ranks.mr_g.age, beside=TRUE) # plot a 'rankogram'
```

```{r eval=FALSE , fig.align="center"}
print(gemtc::sucra(ranks.mr_g.age), digits=3)
```



<br><br>

Here are comparative forest plots comparing treatment effects against one another. I have fixed the gestational age to 26 weeks for illustrative purposes but the range of gestational ages is shown in the covariate plots below these forest plots. 

<br>

```{r, fig.width=8, fig.height=14,fig.align="center",fig.asp=0.30}

weeks <- 26        ### change this value to see effect on forest plots below
gemtc::forest(relative.effect(mcmc.mr_g.age, t1 = "Laser", covariate = weeks))
gemtc::forest(relative.effect(mcmc.mr_g.age, t1 = "Ranibizumab", covariate = weeks))
gemtc::forest(relative.effect(mcmc.mr_g.age, t1 = "Aflibercept", covariate = weeks))
gemtc::forest(relative.effect(mcmc.mr_g.age, t1 = "Bevacizumab", covariate = weeks))
```


<br>

## Covariate plots for gestational age

<br>

Covariate plots below are in log(RR) units.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_g.age, t1="Bevacizumab", t2="Ranibizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>

From the above covariate plot we can see there are no statistically significant differences between Avastin and Lucentis outcomes across gestational ages (credible intervals always incorporate zero).


There is a downward slope suggesting Lucentis may be associated with greater risk of requiring retreatment at lower GA (on the threshold for significance but not quite significant) but then improves to be associated with lower risk at higher GA, although none of these associations reach statistical significance unfortunately.


<br><br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_g.age, t1="Laser", t2="Ranibizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>


*** 


UPDATE (31/12/22) - Between 25.6 and 26.8 weeks GA, Laser is associated with a statistically significant reduction in risk of requiring retreatment when compared to Lucentis. 


***


<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_g.age, t1="Laser", t2="Bevacizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>


*** 


UPDATE (31/12/22) - Again we now have no statististical differences here at any gestational age 


***


<br><br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_g.age, t1="Laser", t2="Aflibercept", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>

All Eylea combination plots look the same as above whereby the are no statistically significant differences between Eylea and all other treatments for any GA. (N.B only Eylea vs Laser shown here but all other Eylea comparisons look very similar). 


<br><br>


***
Summary


*** 

UPDATE (31/12/22) - Between 25.6 and 26.8 weeks GA, Laser is associated with a statistically significant reduction in risk of requiring retreatment when compared to Lucentis. No moderating effect of g.age found on any other treatment modality comparison.

***


***


<br><br>















<br>

### Birth weight as a regressor in the Bayesian meta-regression (potential moderator)

<br>

```{r echo=FALSE}

df_long2_bw <- na.omit(df_long2[,c(1:4,6)])


#### have to remove study Riazi-esfahani 2021 as BW details for Avastin is missing so won't run
#df_long2_bw <- df_long2_bw[!(row.names(df_long2_bw) %in% c(61)) , ]



network.mr_bw <- mtc.network(df_long2_bw, 
                          studies = df_long2_bw[c(1,2,5)],
                          description = "Bayesian NMA")

```


```{r}

set.seed(42)

model.mr_bw <- mtc.model(network.mr_bw, 
                      likelihood = "binom", 
                      link = "log", 
                      linearModel = "random", 
                      n.chain = 4, 
                      type = "regression", 
                      regressor = list(coefficient="unrelated",variable="b.weight", control="Laser"))

#################################################################
# By default, the model will have 4 chains - generate a seed for each - see gemtc documentation
seeds <- sample.int(4, n = .Machine$integer.max)
# Apply JAGS RNG settings to each chain
model.mr_bw$inits <- mapply(c, model.mr_bw$inits, list(
list(.RNG.name="base::Wichmann-Hill", .RNG.seed=seeds[1]),
list(.RNG.name="base::Marsaglia-Multicarry", .RNG.seed=seeds[2]),
list(.RNG.name="base::Super-Duper", .RNG.seed=seeds[3]),
list(.RNG.name="base::Mersenne-Twister", .RNG.seed=seeds[4])), SIMPLIFY=FALSE)
#################################################################

mcmc.mr_bw <- mtc.run(model.mr_bw, 
                n.adapt = 8000, 
                n.iter = 200000, 
                thin = 10)
summary(mcmc.mr_bw)
```


```{r,fig.width=8, fig.height=12,fig.align="center"}
plot(mcmc.mr_bw)
```


```{r, fig.height=6, fig.width=6}
gelman.plot(mcmc.mr_bw)
gelman.diag(mcmc.mr_bw, multivariate = F)
PSRF <- gelman.diag(mcmc.mr_bw)$mpsrf
print(paste0("PSRF = ", round(PSRF, digits = 4)))
```

```{r,fig.align="center"}
ranks.mr_bw <- rank.probability(mcmc.mr_bw, preferredDirection = -1)
rankquant.mr_bw <- rank.quantiles(ranks.mr_bw, probs=c("2.5%"=0.025, "50%"=0.5, "97.5%"=0.975))
```

```{r eval=FALSE, fig.align="center", fig.cap="A"}
plot(ranks.mr_bw, beside=TRUE) # plot a 'rankogram'
```

```{r eval=FALSE , fig.align="center"}
print(gemtc::sucra(ranks.mr_bw), digits=3)
```



<br><br>

Here are comparative forest plots comparing treatment effects against one another. I have fixed the birth weight to 1000g for illustrative purposes. 

<br>

```{r, fig.width=8, fig.height=14,fig.align="center",fig.asp=0.30}

weight <- 1000   ### change value to see effect on foret plots below
gemtc::forest(relative.effect(mcmc.mr_bw, t1 = "Laser", covariate = weight))
gemtc::forest(relative.effect(mcmc.mr_bw, t1 = "Ranibizumab", covariate = weight))
gemtc::forest(relative.effect(mcmc.mr_bw, t1 = "Aflibercept", covariate = weight))
gemtc::forest(relative.effect(mcmc.mr_bw, t1 = "Bevacizumab", covariate = weight))
```


<br>

## Covariate plots for birth weight

<br>

Covariate plots below are in log(RR) units.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_bw, t1="Bevacizumab", t2="Ranibizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

There are no significant differences between treatment outcomes in terms of retreatment rates between Avastin and Lucentis across various birth weights.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_bw, t1="Laser", t2="Ranibizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```


<br>


*** 


UPDATE (31/12/22) - We have a significant moderating effect where birth weights between 820g and 953g are associated with significant reduction in risk of requiring retreatment if treated with Laser vs Lucentis


***


<br><br>



```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_bw, t1="Laser", t2="Bevacizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

There are no significant differences between treatment outcomes in terms of retreatment rates between Avastin and Laser across various birth weights.

<br><br>



```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_bw, t1="Laser", t2="Aflibercept", xlim=NULL, ylim=NULL, 
                    ask=dev.interactive(orNone=TRUE))

```

<br>

Again, there are no statistically significant differences in treatment outcomes in terms of retreatment rates evident between Eylea and any other treatment modality.

<br><br>
















<br><br>

### Mean postmenstrual age at treatment - PMA at treatment (weeks) as a regressor in the Bayesian meta-regression (potential moderator)

<br>


```{r}
df_long2_pma <- na.omit(df_long2[,c(1:4,7)])


network.mr_pma <- mtc.network(df_long2_pma, 
                          studies = df_long2_pma[c(1,2,5)],
                          description = "Bayesian NMA")

```


```{r}

set.seed(42)

model.mr_pma <- mtc.model(network.mr_pma, 
                      likelihood = "binom", 
                      link = "log", 
                      linearModel = "random", 
                      n.chain = 4, 
                      type = "regression", 
                      regressor = list(coefficient="unrelated",variable="pma", control="Laser"))

#################################################################
# By default, the model will have 4 chains - generate a seed for each - see gemtc documentation
seeds <- sample.int(4, n = .Machine$integer.max)
# Apply JAGS RNG settings to each chain
model.mr_pma$inits <- mapply(c, model.mr_pma$inits, list(
list(.RNG.name="base::Wichmann-Hill", .RNG.seed=seeds[1]),
list(.RNG.name="base::Marsaglia-Multicarry", .RNG.seed=seeds[2]),
list(.RNG.name="base::Super-Duper", .RNG.seed=seeds[3]),
list(.RNG.name="base::Mersenne-Twister", .RNG.seed=seeds[4])), SIMPLIFY=FALSE)
#################################################################

mcmc.mr_pma <- mtc.run(model.mr_pma, 
                n.adapt = 8000, 
                n.iter = 200000, 
                thin = 10)
summary(mcmc.mr_pma)
```

```{r,fig.width=8, fig.height=12,fig.align="center"}
plot(mcmc.mr_pma)
```

```{r, fig.height=6, fig.width=6}
gelman.plot(mcmc.mr_pma)
gelman.diag(mcmc.mr_pma, multivariate = F)
PSRF <- gelman.diag(mcmc.mr_pma)$mpsrf
print(paste0("PSRF = ", round(PSRF, digits = 4)))
```

```{r,fig.align="center"}
ranks.mr_pma <- rank.probability(mcmc.mr_pma, preferredDirection = -1)
rankquant.mr_pma <- rank.quantiles(ranks.mr_pma, probs=c("2.5%"=0.025, "50%"=0.5, "97.5%"=0.975))
```

```{r eval=FALSE, fig.align="center", fig.cap="A"}
plot(ranks.mr_pma, beside=TRUE) # plot a 'rankogram'
```

```{r eval=FALSE , fig.align="center"}
print(gemtc::sucra(ranks.mr_pma), digits=3)
```



<br><br>

Here are comparative forest plots comparing treatment effects against one another. I have fixed the pma to 35 weeks for illustrative purposes. 

<br>

```{r, fig.width=8, fig.height=14,fig.align="center",fig.asp=0.30}

pma <- 35.6   ### change value to see effect on forest plots below
gemtc::forest(relative.effect(mcmc.mr_pma, t1 = "Laser", covariate = pma))
gemtc::forest(relative.effect(mcmc.mr_pma, t1 = "Ranibizumab", covariate = pma))
gemtc::forest(relative.effect(mcmc.mr_pma, t1 = "Aflibercept", covariate = pma))
gemtc::forest(relative.effect(mcmc.mr_pma, t1 = "Bevacizumab", covariate = pma))
```


<br>

## Covariate plots for pma (weeks)

<br>

Covariate plots below are in log(RR) units.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_pma, t1="Bevacizumab", t2="Ranibizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

There are no significant differences between treatment outcomes in terms of retreatment rates between Avastin and Lucentis across various pma at treatment.

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_pma, t1="Laser", t2="Ranibizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>




*** 


We have a statististical differences between Lucentis and Laser when taking mean PMA at treatment into account as a moderator. At low PMA (upto 35.6 weeks) treatment with Laser is associated with a statistically significant reduction in risk of requiring retreatment than Lucentis. Beyond 35.6 weeks PMA there are no statistical differences.


***


<br><br>



```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_pma, t1="Laser", t2="Bevacizumab", xlim=NULL, ylim=NULL,
                    ask=dev.interactive(orNone=TRUE))

```

<br>

There are no significant differences between treatment outcomes in terms of retreatment rates between Avastin and Laser across various pma's

<br>


```{r, fig.align="center"}
plotCovariateEffect(mcmc.mr_pma, t1="Laser", t2="Aflibercept", xlim=NULL, ylim=NULL, 
                    ask=dev.interactive(orNone=TRUE))

```





<br><br>


*** 


To summarize

At low PMA (upto 35.6 weeks) treatment with Laser is associated with a statistically significant reduction in risk of requiring retreatment than Lucentis.No other differences are seen at higher PMA or with any other treatment comparisons. 


***



